---
title: "SQL"
output: html_document
---

## Filtering Rows and Columns

With SQL you can filter columns with SELECT and rows with WHERE. Lets look at an example using the Lahman baseball database. First, we need to load some packages.

```{r warning=FALSE, message=FALSE}
library(Lahman)
library(sqldf)
```


Suppose we want to see the homerun totals for the 1927 Yankees. We could write the following:


```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='NYA'and yearID=1927"
sqldf(query)
```


This query identifies players on the NY Yankees team who have hit 40 or more homeruns and had less than 60 strike outs.


```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE teamID='NYA' and HR>=40 and SO<60"
sqldf(query)
```


This query broadens the search to players, on any team in the MLB, who have more than 40 homeruns and less than 60 strikeouts.


```{r}
query<-"SELECT playerID,teamID,yearID,HR,SO FROM Batting
WHERE HR>40 and SO<60"
sqldf(query)
```


This query retrains the search to players on the Philadelphia Phillies team who have more than 30 homeruns during the 70s.


```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE teamID='PHI' and HR>30 and yearID<=1979 and yearID>=1970"
sqldf(query)
```

#------------------------------------------------------------------------

## Ordering Data Outputs

This query again widens the search to players, in desceding order, who have more than 50 homeruns.


```{r}
query<-"SELECT playerID, yearID, teamID,HR FROM Batting
WHERE HR>50
ORDER BY HR DESC"
sqldf(query)
```


This query identified players, on any team in the MLB, that have had a season with at least 400 at bats and less than 10 strikeouts. In this case, it is in ascending order from those with the least strikeouts to the most.


```{r}
query<-"SELECT playerID,yearID,teamID,SO,AB FROM Batting
WHERE SO<10 and AB>=400
ORDER BY SO"
sqldf(query)
```

#------------------------------------------------------------------------


##Aggregation

This query determines the sum of career home runs of Babe Ruth.

```{r}
query<-"SELECT playerID, sum(HR) FROM Batting
WHERE playerID= 'ruthba01'
GROUP BY playerID"
sqldf(query)
```

This query identifies players on any team who have a sum of 600 or less homeruns, specifically ordered in descending order.

```{r}
query<-"SELECT playerID, sum(HR) From Batting
GROUP BY playerID
HAVING sum(HR)>600
ORDER BY sum(HR) DESC"
sqldf(query)
```

In descending order, this query lists players, from any teams, with an average of 30 or less homeruns.

```{r}
query<-"SELECT playerID, avg(HR) From Batting
GROUP BY playerID
HAVING avg(HR)>30
ORDER BY avg(HR) DESC"
sqldf(query)
```

#------------------------------------------------------------------------


## Joins

The joining of tables is a function that allows us to combine data. Below, is just one example of this where we are looking at the Batting table with the Master table. Specifically, we are looking at Babe Ruth in the Batting table.

```{r}
query<-"SELECT nameFirst,nameLast,yearID,teamID,HR 
FROM Batting INNER JOIN Master
ON BATTING.playerID=Master.playerID
WHERE Batting.playerID= 'ruthba01'"
sqldf(query)
```

This next query we are, again, combining two tables to determine players full names in which their seasons had 50 or more homeruns. This is done in decending order. 

```{r}
query<-"SELECT nameFirst,nameLast,yearID,teamID,HR
FROM Batting INNER JOIN Master
ON BATTING.playerID=Master.playerID
WHERE Batting.HR>50
ORDER BY HR DESC"
sqldf(query)
```

Next, we are going to do a query to determine the names of the teams that Babe Ruth his past 22 seasons. It also shows how many homeruns he got each season.


```{r}
query<-"SELECT playerID,name,Batting.yearID,Batting.HR
FROM Batting INNER JOIN Teams
ON Batting.teamID=Teams.teamID AND Batting.yearID=Teams.yearID
WHERE playerID='ruthba01'"
sqldf(query)
```

Next, we are going to move onto our last query. Like with the pervious examples, we are one again merging two tables: the Salary and Master table. In this query we are identifying players with the first name "Bob" who also had an average salary of one million dollars per each year. 

```{r}
query<-"SELECT nameFirst,nameLast,avg(salary)
FROM Salaries INNER JOIN Master
ON Salaries.playerID=Master.playerID
WHERE nameFirst='Bob'
GROUP BY Salaries.playerID
HAVING avg(salary)>1000000
ORDER BY avg(salary) DESC"
sqldf(query)
```


